---
title: "cp_2_processing"
author: "Bonnie EJ Maven"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# 0. Overview

This script is to process the day 8 CP Seurat object output from the first script (1_preprocessing_cp.Rmd). We will remove non-cardiac cells, clean up the data more, then split the data into the desired objects for further analyses in later scripts.

# 1. Set up; Load Libraries & Set Working Directory
Load required libraries for analyses.

```{r}
# Load libraries.
library(Seurat)
library(harmony)
library(ggplot2)

# Set working directory based on the path of the current open file.
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
```

# 2. Set up; Load Seurat Object

```{r}
# Assign a base name for folder and file naming consistency.
baseName = "cp_all"

# Load the seurat_dataset object.
cp_all <- readRDS(file = paste("../data/RDS/", baseName, "_bc.rds", sep = ""))
```

# 3. Identify Cardiac Cells
These markers will identify which clusters have cardiac cells, in addition to which clusters are commonly known contaminant cells in in vitro cardiac differentiations.

```{r}
# Dot plot for general cardiac markers.
general_cardiac <- c("ACTN2", "MYL2", "MYL3", "MYH6", "MYH7", "TNNT2", "TNNI3", "NPPA", "NPPB", "GJA1", "HAND1", "HAND2", "TBX5", "TBX20", "HEY2", "MESP1", "MESP2", "NKX2-5", "FOXF1")

png(paste("../analyses/QC/", baseName, "_DotPlot_cardiac.png", sep = ""), width = 2000, height = 2000, res = 300)
DotPlot(cp_all, features = general_cardiac, group.by = "seurat_clusters")+ theme(axis.text.x = element_text(angle = 45, hjust=1))
dev.off()

# Dot plot for endothelial markers.
endothelial <- c("VWF", "PECAM1", "ESM1", "CD34", "CDH5", "ENG")

png(paste("../analyses/QC/", baseName, "_DotPlot_endothelial.png", sep = ""), width = 2000, height = 2000, res = 300)
DotPlot(cp_all, features = endothelial, group.by = "seurat_clusters")
dev.off()

# Dot plot for endoderm markers.
endoderm <- c("SOX17", "FOXA2", "AFP", "HNF4A", "HHEX")

png(paste("../analyses/QC/", baseName, "_DotPlot_endoderm.png", sep = ""), width = 2000, height = 2000, res = 300)
DotPlot(cp_all, features = endoderm, group.by = "seurat_clusters")
dev.off()

# Dot plot for smooth muscle markers.
smooth_muscle <- c("MYH11", "TAGLN", "CNN1", "ACTA2", "MYLK", "CALD1")

png(paste("../analyses/QC/", baseName, "_DotPlot_smooth_muscle.png", sep = ""), width = 2000, height = 2000, res = 300)
DotPlot(cp_all, features = smooth_muscle, group.by = "seurat_clusters")
dev.off()

# Based on these data cardiac clusters are: 0, 1, 2, 3, 4, 6, 11, 12

# Clean up the environment.
rm(baseName)
```

# 4. Remove Non-Cardiac Clusters
Remove the non-cardiac clusters.

```{r}
# Assign baseName for naming.
baseName2 = "cp_cardiac_all"

# Subset cardiac clusters.
cp_all_cardiac <- subset(cp_all, idents = c("0","1","2","3","4","6","11","12"))

# Due to further downstream issues of the SOX2+ cells, it is clear they are also a contaminating cell type.
cp_all_cardiac_SOX2_0.cells <- WhichCells(object = cp_all_cardiac, expression = SOX2 == 0)
cp_all_cardiac <- subset(cp_all_cardiac, cells = cp_all_cardiac_SOX2_0.cells)

# Run harmony again to account for any additional batch effects.
cp_all_cardiac <- RunHarmony(cp_all_cardiac, group.by.vars = "gem.group", plot_convergence = T, assay.use = "SCT")

# Run Elbowplot to see what PC is the best cut off for the dimensions.
cp_all_cardiac <- RunPCA(cp_all_cardiac, verbose = TRUE)
ElbowPlot(cp_all_cardiac) # Use 30 dimensions.

# Run UMAP and clustering. Use PC # determined from ElbowPlot.
cp_all_cardiac <- RunUMAP(cp_all_cardiac, dims = 1:30, verbose = TRUE)
cp_all_cardiac <- FindNeighbors(cp_all_cardiac, dims = 1:30, verbose = TRUE)
cp_all_cardiac <- FindClusters(cp_all_cardiac, verbose = TRUE, resolution = 0.3)

# Save UMAP as part of QC.
pdf(paste("../analyses/QC/", baseName2, "_bc_DimPlot.pdf", sep = ""))
DimPlot(cp_all_cardiac, reduction = "umap", label = T)
DimPlot(cp_all_cardiac, reduction = "umap", group.by = "gem.group")
dev.off()

# Save Seurat object before proceeding.
saveRDS(cp_all_cardiac, file = paste("../data/RDS/", baseName2, "_bc.rds", sep = ""))

# Clean up the environment.
rm(baseName2)
```

# 5. Subset siRNA Clusters

Now subsetting siRNA Clusters to further analyze in a different script.

```{r}
# Subset out siRNA conditions.
Idents(cp_all_cardiac) <- "gem.group"
table(cp_all_cardiac@meta.data$gem.group, cp_all_cardiac@meta.data$orig.ident)

cp_all_cardiac@meta.data$gem.group
cp_cardiac_siRNA <- subset(cp_all_cardiac, idents = c("0_siRNA_control", "0_siRNA_ISL1"))

# Confirm subset.
table(cp_cardiac_siRNA@meta.data$gem.group)

# Assign baseName for naming.
baseName3 = "cp_cardiac_siRNA"

# Run harmony again to account for any additional batch effects.
cp_cardiac_siRNA <- RunHarmony(cp_cardiac_siRNA, group.by.vars = "gem.group", plot_convergence = T, assay.use = "SCT")

# Run Elbowplot to see what PC is the best cut off for the dimensions.
cp_cardiac_siRNA <- RunPCA(cp_cardiac_siRNA, verbose = TRUE)
ElbowPlot(cp_cardiac_siRNA)
# Use 30 dimensions.

# Run UMAP and clustering. Use PC # determined from ElbowPlot.
cp_cardiac_siRNA <- RunUMAP(cp_cardiac_siRNA, dims = 1:30, verbose = TRUE)
cp_cardiac_siRNA <- FindNeighbors(cp_cardiac_siRNA, dims = 1:30, verbose = TRUE)
cp_cardiac_siRNA <- FindClusters(cp_cardiac_siRNA, verbose = TRUE, resolution = 0.15) # Low resolution.

# Save UMAP as part of QC.
pdf(paste("../analyses/QC/", baseName3, "_bc_DimPlot.pdf", sep = ""))
DimPlot(cp_cardiac_siRNA, reduction = "umap", label = T)
DimPlot(cp_cardiac_siRNA, reduction = "umap", group.by = "gem.group")
dev.off()


# Save Seurat object.
saveRDS(cp_cardiac_siRNA, file = paste("../data/RDS/", baseName3, "_bc.rds", sep = ""))

# Clean up the environment.
rm(baseName3)
```

# 6. Add Genotype Metadata

Adding replicate information to each replicate for day 8 WT, ISL1-/- and NKX2.5-/- cells. Then will recluster and save this object to be analyzed later.

```{r}
# Remove siRNA conditions.
Idents(cp_all_cardiac) <- "gem.group"

table(cp_all_cardiac@meta.data$gem.group)
cp_all_cardiac <- RenameIdents(object = cp_all_cardiac, "0_siRNA_control" = "siRNA_control", "0_siRNA_ISL1" = "siRNA_ISL1")

cp_cardiac_WT_KOs <- subset(cp_all_cardiac, idents = c("siRNA_control", "siRNA_ISL1"), invert = TRUE)

# Label genotype in metadata. IKO, NKO, WT.
cp_cardiac_WT_KOs@meta.data$genotype <- sub("_.*", "", cp_cardiac_WT_KOs@meta.data$gem.group)

# Confirm change.
table(cp_cardiac_WT_KOs@meta.data$gem.group, cp_cardiac_WT_KOs@meta.data$genotype)

# Assign baseName for naming.
baseName4 = "cp_cardiac_WT_KOs"

# Run harmony again to account for any additional batch effects.
cp_cardiac_WT_KOs <- RunHarmony(cp_cardiac_WT_KOs, group.by.vars = "gem.group", plot_convergence = T, assay.use = "SCT")

# Run Elbowplot to see what PC is the best cut off for the dimensions.
cp_cardiac_WT_KOs <- RunPCA(cp_cardiac_WT_KOs, verbose = TRUE)
ElbowPlot(cp_cardiac_WT_KOs)
# Use 15 dimensions.

# Run UMAP and clustering. Use PC # determined from ElbowPlot.
cp_cardiac_WT_KOs <- RunUMAP(cp_cardiac_WT_KOs, dims = 1:15, verbose = TRUE)
cp_cardiac_WT_KOs <- FindNeighbors(cp_cardiac_WT_KOs, dims = 1:15, verbose = TRUE)
cp_cardiac_WT_KOs <- FindClusters(cp_cardiac_WT_KOs, verbose = TRUE)

# Save UMAP as part of QC.
pdf(paste("../analyses/QC/", baseName4, "_bc_DimPlot.pdf", sep = ""))
DimPlot(cp_cardiac_WT_KOs, reduction = "umap", label = T)
DimPlot(cp_cardiac_WT_KOs, reduction = "umap", group.by = "gem.group")
dev.off()

# Save Seurat object.
saveRDS(cp_cardiac_WT_KOs, file = paste("../data/RDS/", baseName4, "_bc.rds", sep = ""))

# Clean up the environment.
rm(baseName4)
```

# 7. Subset WT + ISL1 KO Cells
This is to subset the WT and ISL1 KO cells from the day 8 CPs to analyze separately from the NKX2.5 KO cells.

```{r}
# Remove NKO genotype
Idents(cp_cardiac_WT_KOs) <- "genotype"
cp_cardiac_WT_IKO <- subset(cp_cardiac_WT_KOs, idents = c("NKO"), invert = TRUE)

# Assign baseName for naming.
baseName5 = "cp_cardiac_WT_IKO"

# Run harmony again to account for any additional batch effects.
cp_cardiac_WT_IKO <- RunHarmony(cp_cardiac_WT_IKO, group.by.vars = "gem.group", plot_convergence = T, assay.use = "SCT")

# Run Elbowplot to see what PC is the best cut off for the dimensions.
cp_cardiac_WT_IKO <- RunPCA(cp_cardiac_WT_IKO, verbose = TRUE)
ElbowPlot(cp_cardiac_WT_IKO)
# Use 30 dimensions.

# Run UMAP and clustering. Use PC # determined from ElbowPlot.
cp_cardiac_WT_IKO <- RunUMAP(cp_cardiac_WT_IKO, dims = 1:30, verbose = TRUE)
cp_cardiac_WT_IKO <- FindNeighbors(cp_cardiac_WT_IKO, dims = 1:30, verbose = TRUE)
cp_cardiac_WT_IKO <- FindClusters(cp_cardiac_WT_IKO, verbose = TRUE, resolution = 0.2)

# Save UMAP as part of QC.
pdf(paste("../analyses/QC/", baseName5, "_bc_DimPlot.pdf", sep = ""))
DimPlot(cp_cardiac_WT_IKO, reduction = "umap", label = T)
DimPlot(cp_cardiac_WT_IKO, reduction = "umap", group.by = "gem.group")
dev.off()

# Save Seurat object.
saveRDS(cp_cardiac_WT_IKO, file = paste("../data/RDS/", baseName5, "_bc.rds", sep = ""))

# Clean up the environment.
rm(baseName5)
```

# 8. Subset WT Cells
This is to subset the WT cells from the day 8 CPs to analyze separately from the KO cells.

```{r}
# Remove IKO genotype
Idents(cp_cardiac_WT_IKO) <- "genotype"
cp_cardiac_WT <- subset(cp_cardiac_WT_IKO, idents = c("IKO"), invert = TRUE)

# Assign baseName for naming.
baseName6 = "cp_cardiac_WT"

# Did not recluster because I wanted to maintain the structure of the UMAP to better compare KO data overlaid with WT only data.

# Save UMAP as part of QC.
pdf(paste("../analyses/QC/", baseName6, "_DimPlot.pdf", sep = ""))
DimPlot(cp_cardiac_WT, reduction = "umap", label = T)
DimPlot(cp_cardiac_WT, reduction = "umap", group.by = "gem.group")
dev.off()

# Save Seurat object. I removed the "bc"/batch correction because I did not due it again on this object (see above).
saveRDS(cp_cardiac_WT, file = paste("../data/RDS/", baseName6, ".rds", sep = ""))

# Clean up the environment.
rm(baseName6)
```

# 9. sessionInfo
Capture the current session information.

```{r}
sessionInfo()

# To comment out a line of text use the keyboard shortcut: Command + Shift + C

# 
# R version 4.2.0 (2022-04-22)
# Platform: x86_64-apple-darwin17.0 (64-bit)
# Running under: macOS Monterey 12.4
# 
# Matrix products: default
# LAPACK: /Library/Frameworks/R.framework/Versions/4.2/Resources/lib/libRlapack.dylib
# 
# locale:
#   [1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8
# 
# attached base packages:
#   [1] grid      stats     graphics  grDevices utils     datasets  methods   base     
# 
# other attached packages:
#   [1] SeuratWrappers_0.3.0  sctransform_0.3.3     shiny_1.7.2           harmony_0.1.0        
# [5] Rcpp_1.0.9            RColorBrewer_1.1-3    circlize_0.4.15       ComplexHeatmap_2.12.0
# [9] cowplot_1.1.1         ggsci_2.9             forcats_0.5.1         stringr_1.4.0        
# [13] purrr_0.3.4           readr_2.1.2           tidyr_1.2.0           tibble_3.1.8         
# [17] ggplot2_3.3.6         tidyverse_1.3.2       patchwork_1.1.1       Seurat_4.1.1         
# [21] sp_1.5-0              SeuratObject_4.1.0    dplyr_1.0.9          
# 
# loaded via a namespace (and not attached):
#   [1] utf8_1.2.2            R.utils_2.12.0        reticulate_1.25      
# [4] tidyselect_1.1.2      htmlwidgets_1.5.4     Rtsne_0.16           
# [7] devtools_2.4.4        munsell_0.5.0         codetools_0.2-18     
# [10] ragg_1.2.2            ica_1.0-3             future_1.27.0        
# [13] miniUI_0.1.1.1        withr_2.5.0           spatstat.random_2.2-0
# [16] colorspace_2.0-3      progressr_0.10.1      rstudioapi_0.13      
# [19] stats4_4.2.0          ROCR_1.0-11           tensor_1.5           
# [22] listenv_0.8.0         labeling_0.4.2        polyclip_1.10-0      
# [25] farver_2.1.1          parallelly_1.32.1     vctrs_0.4.1          
# [28] generics_0.1.3        R6_2.5.1              doParallel_1.0.17    
# [31] clue_0.3-61           rsvd_1.0.5            spatstat.utils_2.3-1 
# [34] cachem_1.0.6          assertthat_0.2.1      promises_1.2.0.1     
# [37] scales_1.2.0          googlesheets4_1.0.0   rgeos_0.5-9          
# [40] gtable_0.3.0          globals_0.15.1        processx_3.7.0       
# [43] goftest_1.2-3         rlang_1.0.4           systemfonts_1.0.4    
# [46] GlobalOptions_0.1.2   splines_4.2.0         lazyeval_0.2.2       
# [49] gargle_1.2.0          spatstat.geom_2.4-0   broom_1.0.0          
# [52] BiocManager_1.30.18   reshape2_1.4.4        abind_1.4-5          
# [55] modelr_0.1.8          backports_1.4.1       httpuv_1.6.5         
# [58] tools_4.2.0           usethis_2.1.6         ellipsis_0.3.2       
# [61] spatstat.core_2.4-4   jquerylib_0.1.4       BiocGenerics_0.42.0  
# [64] sessioninfo_1.2.2     ggridges_0.5.3        plyr_1.8.7           
# [67] ps_1.7.1              prettyunits_1.1.1     rpart_4.1.16         
# [70] deldir_1.0-6          pbapply_1.5-0         GetoptLong_1.0.5     
# [73] urlchecker_1.0.1      S4Vectors_0.34.0      zoo_1.8-10           
# [76] haven_2.5.0           ggrepel_0.9.1         cluster_2.1.3        
# [79] fs_1.5.2              magrittr_2.0.3        RSpectra_0.16-1      
# [82] data.table_1.14.2     scattermore_0.8       lmtest_0.9-40        
# [85] reprex_2.0.1          RANN_2.6.1            googledrive_2.0.0    
# [88] fitdistrplus_1.1-8    matrixStats_0.62.0    pkgload_1.3.0        
# [91] hms_1.1.1             mime_0.12             xtable_1.8-4         
# [94] readxl_1.4.0          IRanges_2.30.0        gridExtra_2.3        
# [97] shape_1.4.6           compiler_4.2.0        KernSmooth_2.23-20   
# [100] crayon_1.5.1          R.oo_1.25.0           htmltools_0.5.3      
# [103] mgcv_1.8-40           later_1.3.0           tzdb_0.3.0           
# [106] lubridate_1.8.0       DBI_1.1.3             dbplyr_2.2.1         
# [109] MASS_7.3-58           Matrix_1.4-1          cli_3.3.0            
# [112] R.methodsS3_1.8.2     parallel_4.2.0        igraph_1.3.4         
# [115] pkgconfig_2.0.3       plotly_4.10.0         spatstat.sparse_2.1-1
# [118] xml2_1.3.3            foreach_1.5.2         bslib_0.4.0          
# [121] rvest_1.0.2           callr_3.7.1           digest_0.6.29        
# [124] RcppAnnoy_0.0.19      spatstat.data_2.2-0   cellranger_1.1.0     
# [127] leiden_0.4.2          uwot_0.1.11           curl_4.3.2           
# [130] rjson_0.2.21          lifecycle_1.0.1       nlme_3.1-158         
# [133] jsonlite_1.8.0        viridisLite_0.4.0     fansi_1.0.3          
# [136] pillar_1.8.0          lattice_0.20-45       fastmap_1.1.0        
# [139] httr_1.4.3            pkgbuild_1.3.1        survival_3.3-1       
# [142] glue_1.6.2            remotes_2.4.2         png_0.1-7            
# [145] iterators_1.0.14      stringi_1.7.8         sass_0.4.2           
# [148] profvis_0.3.7         textshaping_0.3.6     memoise_2.0.1        
# [151] irlba_2.3.5           future.apply_1.9.0   
```
