---
title: "cp_5_siRNA_analysis"
author: "Bonnie EJ Maven"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# 0. Overview
This script is to analyze the siRNA conditions (ISL1 knockdown and control knockdown) in day 8 CPs. These data are for Supp Fig 1.

# 1. Set up; Load Libraries & Set Working Directory
Load required libraries for analyses.

```{r}
# Load libraries.
library(Seurat)
library(harmony)
library(ggplot2)
library(monocle3)

# Set working directory based on the path of the current open file.
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
```

# 2. Set up; Load Seurat Object

```{r}
# Assign a base name for folder and file naming consistency.
baseName = "cp_cardiac_siRNA_bc"

dir.create(paste("../analyses/", baseName, sep = ""))

# Load the seurat_dataset object.
cp_cardiac_siRNA <- readRDS(file = paste("../data/RDS/", baseName, ".rds", sep = ""))
```

# 3. Overview of Data

```{r}
table(cp_cardiac_siRNA@meta.data$gem.group)
# siRNA_control    siRNA_ISL1 
# 600              318 

# Overview figures of siRNA-treated cells.
DimPlot(cp_cardiac_siRNA, reduction = "umap", group.by = "gem.group")
DimPlot(cp_cardiac_siRNA, reduction = "umap", label = T, group.by = "seurat_clusters")

# Overview of cardiac genes.
FeaturePlot(cp_cardiac_siRNA,  features = "TNNT2")
FeaturePlot(cp_cardiac_siRNA,  features = "MEF2C")
FeaturePlot(cp_cardiac_siRNA,  features = "NR2F1")
```

# 4. Marker Overview of Clusters
Get markers associated to each cluster (ISL1 siRNA or Control siRNA).

```{r}
genotype_markers <- FindMarkers(object = cp_cardiac_siRNA, group.by = "gem.group", ident.1="siRNA_ISL1", ident.2 = "siRNA_control",  logfc.threshold = 0)
write.csv(genotype_markers, file = paste("../analyses/", baseName, "genotype_markers.csv", sep = ""))

rm(genotype_markers)
```

# 5. Figures for Genes Based on Markers
Plots for Supp Figure 1.

```{r}
# Overview images.
pdf(paste("../analyses", baseName, "_DimPlot_overview.pdf", sep = ""))
DimPlot(cp_cardiac_siRNA, reduction = "umap", group.by = "gem.group")
DimPlot(cp_cardiac_siRNA, reduction = "umap", group.by = "seurat_clusters")
dev.off()

# Early cardiac marker.
pdf(paste("../analyses", baseName, "_VlnPlot_HEY1.pdf", sep = ""))
VlnPlot(cp_cardiac_siRNA,  features = "HEY1", group.by = "gem.group")
dev.off()

# Late cardiac/structural markers displayed with VlnPlot.
pdf(paste("../analyses", baseName, "_VlnPlot_cardiac_markers.pdf", sep = ""))
VlnPlot(cp_cardiac_siRNA,  features = "ACTA2", group.by = "gem.group")
VlnPlot(cp_cardiac_siRNA,  features = "TNNT2", group.by = "gem.group")
VlnPlot(cp_cardiac_siRNA,  features = "NR2F1", group.by = "gem.group")
VlnPlot(cp_cardiac_siRNA,  features = "MYL9", group.by = "gem.group")
dev.off()

# Other cardiac markers displayed with FeaturePlot.
pdf(paste("../analyses", baseName, "_FeaturePlot_cardiac_markers.pdf", sep = ""))
FeaturePlot(cp_cardiac_siRNA, features = "TBX1") # early
FeaturePlot(cp_cardiac_siRNA, features = "HEY1") # early
FeaturePlot(cp_cardiac_siRNA, features = "NKX2-5") # mid to late
FeaturePlot(cp_cardiac_siRNA, features = "TNNT2") # mid to late
FeaturePlot(cp_cardiac_siRNA, features = "MYH6") # structural
FeaturePlot(cp_cardiac_siRNA, features = "NR2F1") # atrial
```

# 6. sessionInfo
Capture all session info current as of these analyses.

```{r}
sessionInfo()

# # 
# R version 4.2.0 (2022-04-22)
# Platform: x86_64-apple-darwin17.0 (64-bit)
# Running under: macOS Monterey 12.5.1
# 
# Matrix products: default
# LAPACK: /Library/Frameworks/R.framework/Versions/4.2/Resources/lib/libRlapack.dylib
# 
# locale:
#   [1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8
# 
# attached base packages:
#   [1] stats     graphics  grDevices utils     datasets  methods   base     
# 
# other attached packages:
#   [1] SeuratWrappers_0.3.0 sctransform_0.3.3    cowplot_1.1.1        ggplot2_3.3.6        harmony_0.1.0       
# [6] Rcpp_1.0.9           Seurat_4.1.1         dplyr_1.0.9          sp_1.5-0             SeuratObject_4.1.0  
# 
# loaded via a namespace (and not attached):
#   [1] Rtsne_0.16            colorspace_2.0-3      deldir_1.0-6          ellipsis_0.3.2        ggridges_0.5.3       
# [6] rstudioapi_0.13       spatstat.data_2.2-0   farver_2.1.1          leiden_0.4.2          listenv_0.8.0        
# [11] remotes_2.4.2         ggrepel_0.9.1         RSpectra_0.16-1       fansi_1.0.3           codetools_0.2-18     
# [16] splines_4.2.0         R.methodsS3_1.8.2     polyclip_1.10-0       jsonlite_1.8.0        ica_1.0-3            
# [21] cluster_2.1.3         png_0.1-7             R.oo_1.25.0           rgeos_0.5-9           uwot_0.1.11          
# [26] shiny_1.7.2           spatstat.sparse_2.1-1 BiocManager_1.30.18   compiler_4.2.0        httr_1.4.3           
# [31] assertthat_0.2.1      Matrix_1.4-1          fastmap_1.1.0         lazyeval_0.2.2        limma_3.52.2         
# [36] cli_3.3.0             later_1.3.0           htmltools_0.5.3       tools_4.2.0           rsvd_1.0.5           
# [41] igraph_1.3.4          gtable_0.3.0          glue_1.6.2            RANN_2.6.1            reshape2_1.4.4       
# [46] scattermore_0.8       vctrs_0.4.1           nlme_3.1-158          progressr_0.10.1      lmtest_0.9-40        
# [51] spatstat.random_2.2-0 stringr_1.4.0         globals_0.15.1        mime_0.12             miniUI_0.1.1.1       
# [56] lifecycle_1.0.1       irlba_2.3.5           goftest_1.2-3         future_1.27.0         MASS_7.3-58          
# [61] zoo_1.8-10            scales_1.2.0          spatstat.core_2.4-4   promises_1.2.0.1      spatstat.utils_2.3-1 
# [66] parallel_4.2.0        RColorBrewer_1.1-3    reticulate_1.25       pbapply_1.5-0         gridExtra_2.3        
# [71] rpart_4.1.16          stringi_1.7.8         rlang_1.0.4           pkgconfig_2.0.3       matrixStats_0.62.0   
# [76] lattice_0.20-45       ROCR_1.0-11           purrr_0.3.4           tensor_1.5            labeling_0.4.2       
# [81] patchwork_1.1.1       htmlwidgets_1.5.4     tidyselect_1.1.2      parallelly_1.32.1     RcppAnnoy_0.0.19     
# [86] plyr_1.8.7            magrittr_2.0.3        R6_2.5.1              generics_0.1.3        DBI_1.1.3            
# [91] pillar_1.8.0          withr_2.5.0           mgcv_1.8-40           fitdistrplus_1.1-8    survival_3.3-1       
# [96] abind_1.4-5           tibble_3.1.8          future.apply_1.9.0    crayon_1.5.1          KernSmooth_2.23-20   
# [101] utf8_1.2.2            spatstat.geom_2.4-0   plotly_4.10.0         grid_4.2.0            data.table_1.14.2    
# [106] digest_0.6.29         xtable_1.8-4          tidyr_1.2.0           httpuv_1.6.5          R.utils_2.12.0       
# [111] munsell_0.5.0         viridisLite_0.4.0    
```
