---
title: "cp_6_WT_KOs_analysis"
author: "Bonnie EJ Maven"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# 0. Overview
This script is to analyze the WT, ISL1 KO and NKX2.5 KO cells from day 8 CPs. These data are for Main Fig 4, Supp Fig 6.

# 1. Set up; Load Libraries & Set Working Directory
Load required libraries for analyses.

```{r}
# Load libraries.
library(Seurat)
library(harmony)
library(ggplot2)
library(monocle3)

# Set working directory based on the path of the current open file.
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
```

# 2. Set up; Load Seurat Object

```{r}
# Assign a base name for folder and file naming consistency.
baseName = "cp_cardiac_WT_KOs_bc"

dir.create(paste("../analyses/", baseName, sep = ""))

# Load the seurat_dataset object.
cp_cardiac_WT_KOs <- readRDS(file = paste("../data/RDS/", baseName, ".rds", sep = ""))
```

# 3. Cluster and Replicate Overview

```{r}
# To make it cleaner, change the order of each genotype so that WT is first.
cp_cardiac_WT_KOs$genotype <- factor(cp_cardiac_WT_KOs$genotype,levels=c("WT", "IKO", "NKO"))

# Overview of clusters; We will save this later on.
DimPlot(object = cp_cardiac_WT_KOs, reduction = "umap", group.by = "seurat_clusters")

# Overview of key cardiac genes.
pdf(paste("../analyses/", baseName, "_FeaturePlot_cardiac_markers.pdf", sep = ""))
FeaturePlot(cp_cardiac_WT_KOs,  features = "TNNT2")
FeaturePlot(cp_cardiac_WT_KOs,  features = "MYL9")
FeaturePlot(cp_cardiac_WT_KOs,  features = "HEY1")
dev.off()

# Overview of genotypes (ISL1 KO, NKX2.5 KO and WT)
pdf(paste("../analyses/", baseName, "_DimPlot_genotypes.pdf", sep = ""))
DimPlot(object = cp_cardiac_WT_KOs, reduction = "umap", group.by = "genotype")
dev.off()


# Overview of replicates (Rep 1, 2 and 3 in each genotype).

# First, subset each genotype.
Idents(cp_cardiac_WT_KOs) <- "genotype"
cp_cardiac_WT_KOs_WTsubset.cells <- WhichCells(object = cp_cardiac_WT_KOs, idents = "WT")
cp_cardiac_WT_KOs_WTsubset <- subset(cp_cardiac_WT_KOs, cells = cp_cardiac_WT_KOs_WTsubset.cells)

cp_cardiac_WT_KOs_IKOsubset.cells <- WhichCells(object = cp_cardiac_WT_KOs, idents = "IKO")
cp_cardiac_WT_KOs_IKOsubset <- subset(cp_cardiac_WT_KOs, cells = cp_cardiac_WT_KOs_IKOsubset.cells)

cp_cardiac_WT_KOs_NKOsubset.cells <- WhichCells(object = cp_cardiac_WT_KOs, idents = "NKO")
cp_cardiac_WT_KOs_NKOsubset <- subset(cp_cardiac_WT_KOs, cells = cp_cardiac_WT_KOs_NKOsubset.cells)

# Now, save these replicate plots.
pdf(paste("../analyses/", baseName, "_DimPlot_replicate_overview.pdf", sep = ""))
DimPlot(cp_cardiac_WT_KOs_WTsubset, reduction = "umap", group.by = "gem.group")
DimPlot(cp_cardiac_WT_KOs_IKOsubset, reduction = "umap", group.by = "gem.group")
DimPlot(cp_cardiac_WT_KOs_NKOsubset, reduction = "umap", group.by = "gem.group")
dev.off()
```

# 4. Marker Overview
Similar analysis to data from Fig 1.

```{r}
# First determine # of cells in each cluster (lowest is 379)
table(cp_cardiac_WT_KOs$seurat_clusters)

# Change ident to seurat_clusters
Idents(cp_cardiac_WT_KOs) <- "seurat_clusters"

# Determine markers for each cluster.
cluster_markers <- FindAllMarkers(cp_cardiac_WT_KOs, only.pos = TRUE, logfc.threshold = 0.25, max.cells.per.ident = 358,  random.seed = 8)

cluster_markers %>%
    group_by(cluster) %>%
    top_n(n = 100, wt = avg_log2FC) -> cluster_markers_top100

write.csv(cluster_markers_top100, paste("../analyses/", baseName, "FindAllMarkers_unnamed_clusters.csv", sep = ""))

# Off of R, each list was analyzed by Panther GO to determine cluster identity. Tentatively, the cluster identity is the following, which will be confirmed with pseudotime analysis.
# 0" = "1_transition", "1" = "3_late", "2" = "0_early", "3" = "4_mixed", "4" = "2_transition")
```

# 5. Monocle Analysis
```{r}
# Confirm correct identity is set.
Idents(cp_cardiac_WT_KOs) <- "seurat_clusters"

# Changing Seurat to CDS format for use in Monocle.
cds <- as.cell_data_set(cp_cardiac_WT_KOs)
cds <- cluster_cells(cds)

# Check how the cells partitioned.
plot_cells(cds, color_cells_by = "partition")

# Plot out the trajectories.
cds <- learn_graph(cds)

# Now choose the starting cell based on marker analyses. This will pop up a new window.
# We chose the starting cell in the original cluster 0 because marker analyses indicated these were early cells.
cds <- order_cells(cds, reduction_method = "UMAP")

# Save this.
pdf(paste("../analyses/", baseName, "_Monocle_labeled.pdf", sep = ""))
plot_cells(cds,
           color_cells_by = "pseudotime",
           label_groups_by_cluster=FALSE,
           label_leaves=FALSE,
           label_branch_points=FALSE, label_cell_groups = FALSE, label_roots = TRUE)
dev.off()

# This confirms our trajectory as we determined by marker analyses. We will now change the Seurat cluster names to reflect this.
```

# 6. Rename Clusters

```{r}
# Save the old idents.
Idents(cp_cardiac_WT_KOs) <- "seurat_clusters"
cp_cardiac_WT_KOs[["original_seurat_clusters"]] <- Idents(object = cp_cardiac_WT_KOs)

# Rename idents according to above work. Namely:
# Cluster 0 will be 1_transition
# Cluster 1 will be 2_late
# Cluster 2 will be 0_early
# Cluster 3 will be 3_mixed
cp_cardiac_WT_KOs <- RenameIdents(object = cp_cardiac_WT_KOs, "0" = "1_transition", "1" = "3_late", "2" = "0_early", "3" = "4_mixed", "4" = "2_transition")

# Save this as a new identity.
cp_cardiac_WT_KOs@meta.data$named_clusters <- cp_cardiac_WT_KOs@active.ident

# Change the order of the identities to make the data cleaner.
cp_cardiac_WT_KOs$named_clusters <- factor(cp_cardiac_WT_KOs$named_clusters,levels=c("0_early", "1_transition", "2_transition", "3_late", "4_mixed"))

# Now rename them so that the transition clusters are the same cluster.
Idents(cp_cardiac_WT_KOs) <- "seurat_clusters"
cp_cardiac_WT_KOs <- RenameIdents(object = cp_cardiac_WT_KOs, "0" = "transition", "1" = "late", "2" = "early", "3" = "mixed", "4" = "transition")

# Save as a new cluster name.
cp_cardiac_WT_KOs@meta.data$named_clusters_2 <- cp_cardiac_WT_KOs@active.ident

# Clean the data.
cp_cardiac_WT_KOs$named_clusters_2 <- factor(cp_cardiac_WT_KOs$named_clusters_2,levels=c("early", "transition", "late", "mixed"))
```

# 7. Overview Figures of Named Clusters
Now that the clusters are properly named, we will generate overview figures for the manuscript.

```{r}
# Frequency of each genotype in cluster.
samp_clus<-table(cp_cardiac_WT_KOs$genotype, cp_cardiac_WT_KOs$named_clusters_2)
samp_clus_plot<-as.data.frame.table(samp_clus)

pdf(paste("../analyses/", baseName, "_ggPlot_cluster_frequency.pdf", sep = ""))
ggplot(samp_clus_plot, aes(fill=Var2, y=Freq, x=Var1)) + geom_bar(position='fill', stat='identity')
dev.off()

# And now save the overview figure.
pdf(paste("../analyses/", baseName, "_DimPlot_named_clusters.pdf", sep = ""))
DimPlot(cp_cardiac_WT_KOs, group.by = "named_clusters") # using original named_clusters here.
dev.off()
```

# 8. Heatmap of Named Clusters
```{r}
# Generate markers again. These should be similar to before, but with the new cluster names integrated into the data
table(cp_cardiac_WT_KOs$named_clusters) # lowest cell number is 445.

Idents(cp_cardiac_WT_KOs) <- "named_clusters_2"

# Now we have to downsample for the cluster analysis and heatmap.
cp_cardiac_WT_KOs_cluster_down.cells <- WhichCells(cp_cardiac_WT_KOs, downsample = 445, seed = 8)
cp_cardiac_WT_KOs_cluster_down <- subset(x = cp_cardiac_WT_KOs, cells = cp_cardiac_WT_KOs_cluster_down.cells)

cluster_markers <- FindAllMarkers(cp_cardiac_WT_KOs_cluster_down, only.pos = TRUE, logfc.threshold = 0.25, random.seed = 8)

cluster_markers %>%
    group_by(cluster) %>%
    top_n(n = 100, wt = avg_log2FC) -> cluster_markers_top100

# Save table.
write.csv(cluster_markers_top100, paste("../analyses/", baseName, "_markers_named_clusters.csv", sep = ""))

# Save heatmap.
png(paste("../analyses/", baseName, "_clusters_top100_deg_heatmap.png", sep = ""), res = 720, height = 20, width = 20)
DoHeatmap(object = cp_cardiac_WT_KOs_cluster_down, features = cluster_markers_top100$gene)
dev.off()
```

# 9. KO Consequences in All Clusters
```{r}
# Determine lowest cell number.
table(cp_cardiac_WT_KOs$named_clusters)

# Downsample to the lowest cell number.
Idents(cp_cardiac_WT_KOs) <- "genotype"
cp_cardiac_WT_KOs_down.cells <- WhichCells(cp_cardiac_WT_KOs_transition_late, downsample = 3174, seed = 8)
cp_cardiac_WT_KOs_down <- subset(cp_cardiac_WT_KOs, cells = cp_cardiac_WT_KOs_down.cells)

# Confirm downsampling.
table(cp_cardiac_WT_KOs_down@meta.data$genotype)

# FindMarkers for each genotype.
Idents(cp_cardiac_WT_KOs_down) <- "genotype"

WT_KOs_down_IKO_markers <- FindMarkers(cp_cardiac_WT_KOs_down, ident.1 = "IKO", ident.2 = "WT")
WT_KOs_down_NKO_markers <- FindMarkers(cp_cardiac_WT_KOs_down, ident.1 = "NKO", ident.2 = "WT")

# Save these tables.
write.csv(WT_KOs_down_IKO_markers, paste("../analyses/", baseName, "FindAllMarkers_allClusters_IKO.csv", sep = ""))
write.csv(WT_KOs_down_NKO_markers, paste("../analyses/", baseName, "FindAllMarkers_allClusters_NKO.csv", sep = ""))
```

# 10. KO Consequences in Transition and Late Clusters
```{r}
# Separating out the transition and late clusters.
Idents(cp_cardiac_WT_KOs) <- "named_clusters_2"
cp_cardiac_WT_KOs_transition_late.cells <- WhichCells(object = cp_cardiac_WT_KOs, idents = c("transition", "late"), seed = 8)
cp_cardiac_WT_KOs_transition_late <- subset(x = cp_cardiac_WT_KOs, cells = cp_cardiac_WT_KOs_transition_late.cells)

# Lowest cell number (3174 cells)
table(cp_cardiac_WT_KOs_transition_late@meta.data$genotype)

# Downsample to the lowest cell number.
Idents(cp_cardiac_WT_KOs_transition_late) <- "genotype"
cp_cardiac_WT_KOs_transition_late_down.cells <- WhichCells(cp_cardiac_WT_KOs_transition_late, downsample = 3174, seed = 8)
cp_cardiac_WT_KOs_transition_late_down <- subset(x = cp_cardiac_WT_KOs_transition_late, cells = cp_cardiac_WT_KOs_transition_late_down.cells)

# Confirm downsampling.
table(cp_cardiac_WT_KOs_transition_late_down@meta.data$genotype)

# FindMarkers for each genotype.
Idents(cp_cardiac_WT_KOs_transition_late_down) <- "genotype"

transition_late_down_IKO_markers <- FindMarkers(cp_cardiac_WT_KOs_transition_late_down, ident.1 = "IKO", ident.2 = "WT")
transition_late_down_NKO_markers <- FindMarkers(cp_cardiac_WT_KOs_transition_late_down, ident.1 = "NKO", ident.2 = "WT")

# Save these tables.
write.csv(transition_late_down_IKO_markers, paste("../analyses/", baseName, "FindAllMarkers_TransitionLate_IKO.csv", sep = ""))
write.csv(transition_late_down_NKO_markers, paste("../analyses/", baseName, "FindAllMarkers_TransitionLate_NKO.csv", sep = ""))

# Heatmap to display these markers.

# Marker list to display based on GO analyses done separately.
up_KOs <- readLines(con = paste("../analyses/", baseName, "up_KOs_input.csv"))
down_KOs <- readLines(con = paste("../analyses/", baseName, "down_KOs_input.csv"))
up_NKO <- readLines(con = paste("../analyses/", baseName, "up_NKO_input.csv"))

transition_late_markers <- c(up_KOs, down_KOs, up_NKO)

DoHeatmap(cp_cardiac_WT_KOs_transition_late_down, features = transition_late_markers)
```


# 11. sessionInfo
Capture all session info current as of these analyses.

```{r}
sessionInfo()

# to comment out a line of text use the keyboard shortcut: Command + Shift + C
# 
# # 
# R version 4.2.0 (2022-04-22)
# Platform: x86_64-apple-darwin17.0 (64-bit)
# Running under: macOS Monterey 12.5.1
# 
# Matrix products: default
# LAPACK: /Library/Frameworks/R.framework/Versions/4.2/Resources/lib/libRlapack.dylib
# 
# locale:
#   [1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8
# 
# attached base packages:
#   [1] stats     graphics  grDevices utils     datasets  methods   base     
# 
# other attached packages:
#   [1] SeuratWrappers_0.3.0 sctransform_0.3.3    cowplot_1.1.1        ggplot2_3.3.6        harmony_0.1.0       
# [6] Rcpp_1.0.9           Seurat_4.1.1         dplyr_1.0.9          sp_1.5-0             SeuratObject_4.1.0  
# 
# loaded via a namespace (and not attached):
#   [1] Rtsne_0.16            colorspace_2.0-3      deldir_1.0-6          ellipsis_0.3.2        ggridges_0.5.3       
# [6] rstudioapi_0.13       spatstat.data_2.2-0   farver_2.1.1          leiden_0.4.2          listenv_0.8.0        
# [11] remotes_2.4.2         ggrepel_0.9.1         RSpectra_0.16-1       fansi_1.0.3           codetools_0.2-18     
# [16] splines_4.2.0         R.methodsS3_1.8.2     polyclip_1.10-0       jsonlite_1.8.0        ica_1.0-3            
# [21] cluster_2.1.3         png_0.1-7             R.oo_1.25.0           rgeos_0.5-9           uwot_0.1.11          
# [26] shiny_1.7.2           spatstat.sparse_2.1-1 BiocManager_1.30.18   compiler_4.2.0        httr_1.4.3           
# [31] assertthat_0.2.1      Matrix_1.4-1          fastmap_1.1.0         lazyeval_0.2.2        limma_3.52.2         
# [36] cli_3.3.0             later_1.3.0           htmltools_0.5.3       tools_4.2.0           rsvd_1.0.5           
# [41] igraph_1.3.4          gtable_0.3.0          glue_1.6.2            RANN_2.6.1            reshape2_1.4.4       
# [46] scattermore_0.8       vctrs_0.4.1           nlme_3.1-158          progressr_0.10.1      lmtest_0.9-40        
# [51] spatstat.random_2.2-0 stringr_1.4.0         globals_0.15.1        mime_0.12             miniUI_0.1.1.1       
# [56] lifecycle_1.0.1       irlba_2.3.5           goftest_1.2-3         future_1.27.0         MASS_7.3-58          
# [61] zoo_1.8-10            scales_1.2.0          spatstat.core_2.4-4   promises_1.2.0.1      spatstat.utils_2.3-1 
# [66] parallel_4.2.0        RColorBrewer_1.1-3    reticulate_1.25       pbapply_1.5-0         gridExtra_2.3        
# [71] rpart_4.1.16          stringi_1.7.8         rlang_1.0.4           pkgconfig_2.0.3       matrixStats_0.62.0   
# [76] lattice_0.20-45       ROCR_1.0-11           purrr_0.3.4           tensor_1.5            labeling_0.4.2       
# [81] patchwork_1.1.1       htmlwidgets_1.5.4     tidyselect_1.1.2      parallelly_1.32.1     RcppAnnoy_0.0.19     
# [86] plyr_1.8.7            magrittr_2.0.3        R6_2.5.1              generics_0.1.3        DBI_1.1.3            
# [91] pillar_1.8.0          withr_2.5.0           mgcv_1.8-40           fitdistrplus_1.1-8    survival_3.3-1       
# [96] abind_1.4-5           tibble_3.1.8          future.apply_1.9.0    crayon_1.5.1          KernSmooth_2.23-20   
# [101] utf8_1.2.2            spatstat.geom_2.4-0   plotly_4.10.0         grid_4.2.0            data.table_1.14.2    
# [106] digest_0.6.29         xtable_1.8-4          tidyr_1.2.0           httpuv_1.6.5          R.utils_2.12.0       
# [111] munsell_0.5.0         viridisLite_0.4.0   
```
